---
title: "assignment"
author: "Lu Zimeng"
date: "`r Sys.Date()`"
output: html_document
---
```{r}
BiocManager::install("ETHZ-INS/epiwraps")
BiocManager::install("rGREAT")
```

```{r}
suppressPackageStartupMessages({
  library(GenomicRanges)
  library(epiwraps)
  library(ggplot2)
  library(rGREAT)
})
```
## Download the data

```{r, eval=FALSE}
options(timeout = 6000)
download.file("https://ethz-ins.org/content/w10.assignment.zip", "w10.assignment.zip")
unzip("w10.assignment.zip")
list.files()
```
## Prepare the regions and the tracks

```{r}
tracks <- list.files(pattern="bw$")
region1 <- rtracklayer::import.bed("Creb1.bed")
region3 <- rtracklayer::import.bed("Creb3.bed")
region3L1 <- rtracklayer::import.bed("Creb3L1.bed")
```

```{r}
regions <- reduce(unlist(GRangesList(region1,region3,region3L1)))
```

## Plot

```{r}
ml <- signal2Matrix(tracks, regions, extend=2000)
set.seed(123)  
nf <- getNormFactors(tracks, useSeqLevels="chr1")
ml2 <- rescaleSignalMatrices(ml, scaleFactors=nf)
plotEnrichedHeatmaps(ml)
```

## Clustering

```{r}
set.seed(123)  # to ensure that it gives the same results everytime
cl <- clusterSignalMatrices(ml, k=4)
# the number of regions assigned to each cluster
table(cl)
# the first few values of the cluster assignment vector
head(cl)
# the total number of regions assigned to clusters
length(cl)
# the total number of genomic regions in the input data
length(regions)
# contains the cluster assignment for each genomic region
regions$cluster <- cl
```

```{r}
mycolors <- c("1"="red", "2"="blue", "3"="darkgreen", "4"="black")
plotEnrichedHeatmaps(ml, row_split=cl, mean_color=mycolors)
# clusters 1&2 are enriched in Creb3, Identifying that Creb3 binding sites are in clusters 1&2, and clusters 1&2 are potentially regulated by Creb3; Creb1 & Creb3L1 are specifically enriched in clusters 2&4, suggesting that they are co-regulated specific biological processes.
```

```{r}
cl2 <- clusterSignalMatrices(ml, k=2:10)
ggplot(cl2$varExplained, aes(k, varExplained)) + geom_line()
# From the "elbow" of the curve, we can conclude that the optimal number of clusters to use in the analysis is 3or4or5
```


```{r}
# Plotting just the averages:
d <- meltSignals(ml, splitBy=cl)
ggplot(d, aes(position, mean, colour=sample)) + geom_line(size=1.2) + facet_wrap(~split)
#  Creb3 is important in defining the unique cellular state or differentiation trajectory represented by cluster 1. Creb3L1& Creb1 are the key transcriptional regulators in cluster 4.
```

# Clustering using relative signal instead:

```{r}
cl <- clusterSignalMatrices(ml, k=4, scaleRows = TRUE)
d <- meltSignals(ml, splitBy=cl)
ggplot(d, aes(position, mean, colour=sample)) + geom_line() + facet_wrap(~split)
plotEnrichedHeatmaps(ml, row_split = cl, scale_rows = "global")
#  co-regulated regulated regions (cluster 1) that are enriched for the binding of TF Creb1 & Creb3L1. Differentially regulated regions( cluster 2 & 3 & 4 )that are enriched for the binding of specific proteins ( Creb1 & 3 & 3L1 sperately).
```
## Enrichment analysis

# Find what's enriched in one cluster with respect to the others:

```{r}
# we first split the regions by cluster:
split_regions <- split(regions, cl)

# we send the data to the GREAT server to run an enrichment analysis
job <- submitGreatJob(gr=split_regions[["4"]], bg=regions, species="hg38")
res <- getEnrichmentTables(job)
names(res)
```
# We plot the top Biological Processes:

```{r, fig.width=9, fig.height=6}
bp <- res$`GO Biological Process`
head(bp)

ggplot(head(bp,15), aes(Hyper_Fold_Enrichment, reorder(name, Hyper_Adjp_BH), size=Hyper_Foreground_Region_Hits, color=-log10(Hyper_Adjp_BH))) + geom_point() + scale_color_viridis_c()
```
# mitochondrial tRNA processing/ modification & RNA processing have the largest the fold-enrichment of the Biological Process in the foreground set compared to the background set, but the numbers of foreground region hits for them are small. 
# The top enriched Biological Processes is tRNA processing, as it has smallest P-value ( with yellow color and order in the bottom).



